Edge sign prediction in general graphs and Correlation Clustering

### 2017-10-03 troll/experiments.tex
corrections in troll related and experiments.

We consider six real-world classification datasets. The first four are \dssn{} widely
used as benchmarks for this
task~(\eg{}\autocites{Leskovec2010}{shahriari2014ranking}{wu2016troll}{OnlineCompletion17}):
\begin{newcontent}
In \adv{}, a trust-based social network for open source developers, a user $u$ can certify another
user $v$ with different degrees of trust: \enquote{Observer}, \enquote{Apprentice} (both of which we
consider negative), \enquote{Journeyer} and \enquote{Master} (both of which we consider
positive).\footnote{We download the 7\thup{} of July, 2014 version from
\url{http://www.trustlet.org/datasets/advogato/}.} A full description of this trust metric, and its
resistance to attacks, is available in the PhD thesis of the website's creator~\autocite[Section
4]{AdvogatoTrustThesis02}).
\end{newcontent}
In \wik{}, there is an edge from user $u$ to user $v$ if $v$ applies for an admin position and $u$
votes for or against that promotion. In \sla{}, a news sharing and commenting website, member $u$
can tag other members $v$ as friends or foes. Finally, in \epi{}, an online shopping website, user
$v$ reviews products and, based on these reviews, another user $u$ can display whether he considers
$v$ to be reliable or not. In addition to these four datasets, we considered two other \ssn{} where
the signs are inferred automatically, rather than given explicitly by the users.  In
\kiw{}~\cite{wikiedits11}, an edge from Wikipedia user $u$ to user $v$ indicates whether they edited
the same article in a constructive manner or not.\footnote{This is the
\href{http://konect.uni-koblenz.de/networks/wikisigned-k2}{KONECT version of the
\enquote{Wikisigned} dataset}, from which we removed self-loops.} Finally, in the
\aut{}~\cite{kumar2016structure} network, an author $u$ cites another author $v$ by either endorsing
or criticizing $v$'s work. The edge sign is derived by classifying the citation sentiment with a
technique using a list of positive and negative words; see
\cite{kumar2016structure} for more details.\footnote{We again removed self-loops and merged
multi-edges which are all of the same sign.}

@@

Although the semantics of the edge signs is not the same across these networks, we can see from
\autoref{tab:dataset} that our generative model essentially fits all of them. Specifically,
two columns of the table report the rate of label (ir)regularity, as measured by
$\Psi^2_{G''}(Y)/|E|$ (6\thup{} column) and $\Psi_{G}(Y)/|E|$ (7\thup{} column), where 
\[
\Psi^2_{G''}(Y) = \min_{(\bp,\bq)} \left(f_{E_0}(\bp,\bq) + f_{E\setminus E_0}(\bp,\bq)\right)\,,
\]
$f_{E_0}$ and $f_{E\setminus E_0}$ being the quadratic criterions of \autoref{ss:passive}, viewed as
functions of both $(\bp,\bq)$, and \yuv{} when all labels are known, and $\Psi_{G}(Y)$ is the
label regularity measure adopted in the online setting, as defined in \autoref{s:prel}. It is
reasonable to expect that higher label irregularity corresponds to lower prediction performance.
This trend is in fact confirmed by our experimental findings: whereas \epi{} tends to be easy,
\aut{} tends to be hard, and this holds for all algorithms we tested, even if they do not explicitly
comply with our inductive bias principles. Moreover, $\Psi^2_{G''}(Y)/|E|$ tends to be proportional
to $\Psi_{G}(Y)/|E|$ across datasets, hence confirming the anticipated connection between the two
regularity measures.

\begin{newcontent}
Finally, there is a low fraction of reciprocal edges (\ie{} both $\euv \in E$ and $\evu
\in E$), which is a common mechanism of link formation in directed
networks~\autocites{DirectedReciprocity04}{Reciprocity13}). Moreover, in most cases, such reciprocal edges do not disagree,
\ie{} they have the same sign. In practice, we can use this fact to improve our accuracy at no
additional computational cost: when predicting $\euv \in E_{\mathrm{test}}$, if the reciprocal edge
\evu{} is part of the training set, we set $\yhat_{u,v} = y_{v,u}$. For clarity, when comparing
our methods with existing approaches, we not use that trick. But afterwards, we show in
\autoref{tab:twin} when it can be beneficial.
% more references in  http://www.nature.com/articles/srep02729
\end{newcontent}

\begin{newcontent}
\subsection{Synthetic signs}

Recall that according to our generative model of \autoref{s:gen}, each node $u$ has a parameter $p_u$
governing its sending behavior and each node $v$ has a parameter $q_v$ governing its receiving
behavior, such that the sign of the edge \euv{} is positive with probability $\frac{p_u+q_v}{2} =
\etauv$. Given a topology $G=(V,E)$, we start by assigning a $p$ and $q$ values to each node. If we
want \etauv{} to be uniform over $E$, we have to take into account the out-degree of nodes with at
least one outgoing edge, and likewise the in-degree of nodes with at least one incoming edge. For
that, in the case of $p$, we partition the interval $[0,1]$ into $|V|$ segments of size proportional
to $\dout(u_1),
\dout(u_2), \ldots, \dout(u_{|V|})$. We shuffle these segments and draw a number \uar{} from each of
them, which we set as the $p$ value of the corresponding node. Then, because we want to model real
sign distribution and have more
positive edges, we apply an exponential transformation \[ p \rightarrow \frac{1}{1-e^{-\lambda}}
\left( 1-e^{-\lambda p} \right) \,,\] where we choose $\lambda=3$. We do the same for $q$, then for
every edge \euv{}, we set it positive with probability \etauv{}. An example of the distributions we
obtain for the \wik{} graph is showed on \autoref{fig:troll_synth_distrib}, giving $71.9\%$ of
positive edges.

\vspace{\baselineskip}
\noindent\begin{minipage}{187mm} 
  \begin{minipage}{46mm}
    \centering
    \includegraphics[width=.95\linewidth]{raw/synth_wik_pdense.pdf}
    \captionof{subfigure}{Density of $p$.}
    % \label{fig:figure1}
  \end{minipage}%
  \begin{minipage}{46mm}
    \centering 
    \includegraphics[width=.95\linewidth]{raw/synth_wik_qdense.pdf} 
    \captionof{subfigure}{Density of $q$.} 
    % \label{fig:figure2} 
  \end{minipage} 
  \begin{minipage}{46mm}
    \centering 
    \includegraphics[width=.95\linewidth]{raw/synth_wik_etadense.pdf} 
    \captionof{subfigure}{Density of $\eta$.} 
    % \label{fig:figure2} 
  \end{minipage} 
  \begin{minipage}{46mm}
    \centering 
    \includegraphics[width=.95\linewidth]{raw/synth_wik_signdense.pdf}
    \captionof{subfigure}{Density of $\eta$ for both signs.} 
    % \label{fig:figure2} 
  \end{minipage} 
  \captionof{figure}{Synthetic distributions on \wik{}.} 
  \label{fig:troll_synth_distrib} 
\end{minipage}
\vspace{\baselineskip}

Once the signs are generated, we select a training set \trainset{} \uar{}, and predict
the signs of the remaining edges in the testing set $E\setminus \trainset{}$. The Bayes optimal
predictor classifies an edge \euv{} as positive if $\etauv \geq \shalf$ and negative otherwise. 
We compare it with the algorithm analyzed in \autoref{ss:bayes_approx}, which we call \usrule{}
(Bayes Learning Classifier based on \emph{tr}ollness and \emph{un}trustworthiness). In practice,
\usrule{} proceeds as follow: after computing $\htr(u)$ and $\hun(u)$ on training set $E_0$ for all
$u \in V$ (or setting those values to \shalf{} in case there is no outgoing or incoming edges for
some node), we use the equation \eqref{eq:predictor}\footnote{We reproduce this equation here for
convenience: $\sgn\Big(\big(1-\htr(u)\big) + \big(1-\hun(v)\big) -\tau - \tfrac{1}{2}\Big)$.},
having estimated $\tau$ on $E_0$.

@@

Finally, we can also look in more details at the role of $p$ and $q$ by building a 2D histogram of
the testing edges \euv{} based on their coordinate $(p_u, q_v)$. For instance on
\autoref{fig:tr_synth_mae}, we see that the MAE between $\eta$ and $\wh{\eta}$ is not uniform over
the $p,q$-space. It seems to decrease along the diagonal from $(0,0)$ to $(1,1)$. Note though that
\autoref{fig:tr_synth_sizemae} suggests this is likely due to the higher number of edges in the top
right corner (reflecting the imbalance toward positive edges), which allows better estimation. 
In this case, we cannot use the MCC because off-diagonal, the Bayes predictor classifies edges
either all positive or all negative, resulting in a division by zero in the definition. We thus
fall back on accuracy and display in \autoref{fig:tr_synth_acc} the difference between the accuracy
of the Bayes predictor and the accuracy of \usrule{}. This time, the anti diagonal where
$\frac{p+q}{2}=\lhalf$ seems to play a special role. Indeed, as showed more clearly in
\autoref{fig:tr_synth_cmpacc}, the gap in accuracy between the Bayes predictor and \usrule{}
increases symmetrically as $\frac{p+q}{2}$ moves away from \shalf{}. The gap also tends to disappear
when $\frac{p+q}{2}=\lhalf$, for in that region, both predictors can only rely on random predictions.

While we show those patterns on \wik{} with a 40\% training size, they are consistent across
datasets and training size.
\end{newcontent}

@@

  \begin{subfigure}[b]{0.504\textwidth} 
    \includegraphics[width=\textwidth]{raw/synthgen_wik_rule_40_acc.pdf} 
    \caption{100 times the accuracy.} 
    \label{fig:tr_synth_acc}
  \end{subfigure}~
  \begin{subfigure}[b]{0.504\textwidth} 
    \includegraphics[width=\textwidth]{raw/synthgen_wik_rule_40_cmpacc.pdf} 
    \caption{Difference of accuracy between \usrule{} and Bayes as $\frac{p+q}{2}$ moves away from $\frac{1}{2}$.} 
    \label{fig:tr_synth_cmpacc}
  \end{subfigure}
  \caption{Results on \wik{} with \usrule{} on a training set of size 40\%}
\end{figure*}

@@

  \item The \usrule{} algorithm, as described above.

@@

Our main results are summarized in \autoref{tab:all_mcc}, reporting MCC test set performance after
training on sets of varying size (from 5\% to 25\%, plus 50\% and 90\%). Results have been averaged
over 12 repetitions.
Because scalability is a major concern when training on sizeable datasets, we also give an idea of
relative training times (in milliseconds) by reporting the time it took to train a single run of
each algorithm on a training set of size\footnote{Comparison of training time performances is fair
since all algorithms have been carefully implemented using the same stack of Python libraries, and
run on the same machine (16 Xeon cores and 192Gb Ram).} 15\% of $|E|$, and then predict on the test
set. Though our experiments are not conclusive, some trends can be readily spotted:

@@

  \item \uslpropGsec{} always ranks first or at least second in this comparison when MCC is
    considered. On top of it, \uslpropGsec{} is fastest among the global methods (one or even two
    orders of magnitude faster), thereby showing the benefit of our approach to edge sign
    prediction.

@@

\begin{newcontent}
\subsection{Additional experiments}

@@

In the second set of experiments, we study the effect of our hypothesis that the labeled edges of the training set are
chosen \uar{}. In two of our datasets (\wik{} and \epi{}), edges come with the timestamp at which they
were created. A more realistic way of choosing the training set is therefore, for a given training
size $m$, to let $E_0$ be the $m$ oldest edges, and try to predict the remaining, newest ones. This
is a common experimental setting in link prediction, where the goal is to infer which pair of nodes
will be connected in the future~\autocite{linkPredSurvey16}. However, as showed in
\autoref{tab:troll_ts_mcc}, this makes the problem much more challenging, both for our methods and
out competitors. It is quite surprising that even when the training set is as large as 90\% of all
edges, the MCC are so low compared with those reported in \autoref{tab:all_mcc}.

@@

@@

\end{newcontent}

>>>>

We consider six real-world classification datasets. The first four are \dssn{} widely
used as benchmarks for this
task~(\eg{}\autocites{Leskovec2010}{shahriari2014ranking}{wu2016troll}{OnlineCompletion17}):
In \adv{}, a trust-based social network for open source developers, a user $u$ can certify another
user $v$ with different degrees of trust: \enquote{Observer}, \enquote{Apprentice} (both of which we
consider negative), \enquote{Journeyer} and \enquote{Master} (both of which we consider
positive).\footnote{We download the 7\thup{} of July, 2014 version from
\url{http://www.trustlet.org/datasets/advogato/}.} A full description of this trust metric, and its
resistance to attacks, is available in the PhD thesis of the website's creator~\autocite[Section
4]{AdvogatoTrustThesis02}).
In \wik{}, there is an edge from user $u$ to user $v$ if $v$ applies for an admin position and $u$
votes for or against that promotion. In \sla{}, a news sharing and commenting website, member $u$
can tag other members $v$ as friends or foes. Finally, in \epi{}, an online shopping website, user
$v$ reviews products and, based on these reviews, another user $u$ can display whether he considers
$v$ to be reliable or not. In addition to these four datasets, we considered two other \ssn{} where
the signs are inferred automatically, rather than given explicitly by the users.  In
\kiw{}~\cite{wikiedits11}, an edge from Wikipedia user $u$ to user $v$ indicates whether they edited
the same article in a constructive manner or not.\footnote{This is the
\href{http://konect.uni-koblenz.de/networks/wikisigned-k2}{KONECT version of the
\enquote{Wikisigned} dataset}, from which we removed self-loops.} Finally, in the
\aut{}~\cite{kumar2016structure} network, an author $u$ cites another author $v$ by either endorsing
or criticizing $v$'s work. The edge sign is derived by classifying the citation sentiment with a
technique using a list of positive and negative words; see
\cite{kumar2016structure} for more details.\footnote{We again removed self-loops and merged
multi-edges which are all of the same sign.}

@@

Although the semantics of the edge signs is not the same across these networks, we can see from
\autoref{tab:dataset} that our generative model essentially fits all of them. Specifically,
two columns of the table report the rate of label (ir)regularity, as measured by
$\Psi^2_{G''}(Y)/|E|$ (6\thup{} column) and $\Psi_{G}(Y)/|E|$ (7\thup{} column), where 
\[
  \Psi^2_{G''}(Y) = \min_{(\bp,\bq)} \sum_{(u,v) \in E}
  \left(\frac{1+\yuv}{2} - \frac{p_u+q_v}{2} \right)^2
\]
as first described in \autoref{ss:passive}
and $\Psi_{G}(Y)$ is the
label regularity measure adopted in the online setting, as defined in \autoref{s:prel}. It is
reasonable to expect that higher label irregularity corresponds to lower prediction performance.
This trend is in fact confirmed by our experimental findings: whereas \epi{} tends to be easy,
\aut{} tends to be hard, and this holds for all algorithms we tested, even if they do not explicitly
comply with our inductive bias principles. Moreover, $\Psi^2_{G''}(Y)/|E|$ tends to be proportional
to $\Psi_{G}(Y)/|E|$ across datasets, hence confirming the anticipated connection between the two
regularity measures.

Finally, there is a low fraction of reciprocal edges (\ie{} both $\euv \in E$ and $\evu
\in E$), which is a common mechanism of link formation in directed
networks~\autocites{DirectedReciprocity04}{Reciprocity13}). Moreover, in most cases, such reciprocal edges do not disagree,
\ie{} they have the same sign. In practice, we can use this fact to improve our accuracy at no
additional computational cost: when predicting $\euv \in E_{\mathrm{test}}$, if the reciprocal edge
\evu{} is part of the training set, we set $\yhat_{u,v} = y_{v,u}$. For clarity, when comparing
our methods with existing approaches, we not use that trick. But afterwards, we show in
\autoref{tab:twin} when it can be beneficial.
% more references in  http://www.nature.com/articles/srep02729

\subsection{Synthetic signs}

Recall that according to our generative model of \autoref{s:gen}, each node $u$ has a parameter $p_u$
governing its sending behavior and each node $v$ has a parameter $q_v$ governing its receiving
behavior, such that the sign of the edge \euv{} is positive with probability $\frac{p_u+q_v}{2} =
\etauv$. Given a topology $G=(V,E)$, we start by assigning a $p$ and $q$ values to each node. If we
want \etauv{} to be uniform over $E$, we have to take into account the out-degree of nodes with at
least one outgoing edge, and likewise the in-degree of nodes with at least one incoming edge. For
that, in the case of $p$, we partition the interval $[0,1]$ into $|V|$ segments of size proportional
to $\dout(u_1),
\dout(u_2), \ldots, \dout(u_{|V|})$. We shuffle these segments and draw a number \uar{} from each of
them, which we set as the initial $p'$ value of the corresponding node. Then, because we want to model real
sign distribution and have more
positive than negative edges, we apply an exponential transformation \[ p = \frac{1}{1-e^{-\lambda}}
\left( 1-e^{-\lambda p'} \right) \,,\] where we choose $\lambda=3$. We do the same for $q$, then for
every edge \euv{}, we set it positive with probability \etauv{}. An example of the distributions we
obtain for the \wik{} graph is showed on \autoref{fig:troll_synth_distrib}\footnote{More precisely,
these figures are normalized histogram, meaning the counts are scale down so that the bars total
area sums up to one.}, giving $71.9\%$ of positive edges.

% TODO: do a proper subfigure instead of minipage?
\vspace{\baselineskip}
\noindent\begin{minipage}{187mm} 
  \begin{minipage}{46mm}
    \centering
    \includegraphics[width=.95\linewidth]{raw/synth_wik_pdense.pdf}
    \captionof{subfigure}{Density of $p$.}
    % \label{fig:figure1}
  \end{minipage}%
  \begin{minipage}{46mm}
    \centering 
    \includegraphics[width=.95\linewidth]{raw/synth_wik_qdense.pdf} 
    \captionof{subfigure}{Density of $q$.} 
    % \label{fig:figure2} 
  \end{minipage} 
  \begin{minipage}{46mm}
    \centering 
    \includegraphics[width=.95\linewidth]{raw/synth_wik_etadense.pdf} 
    \captionof{subfigure}{Density of $\eta$.} 
    % \label{fig:figure2} 
  \end{minipage} 
  \begin{minipage}{46mm}
    \centering 
    \includegraphics[width=.95\linewidth]{raw/synth_wik_signdense.pdf}
    \captionof{subfigure}{Density of $\eta$ for both signs.} 
    % \label{fig:figure2} 
  \end{minipage} 
  \captionof{figure}{Synthetic distributions on \wik{}.} 
  \label{fig:troll_synth_distrib} 
\end{minipage}
\vspace{\baselineskip}

Once the signs are generated, we select a training set \trainset{} \uar{}, and predict
the signs of the remaining edges in the testing set $E\setminus \trainset{}$. The Bayes optimal
predictor classifies an edge \euv{} as positive if $\etauv \geq \shalf$ and negative otherwise. 
We compare it with the \usrule{} algorithm analyzed in \autoref{ss:bayes_approx}. Recall that
\usrule{} proceeds as follow: after computing $\htr(u)$ and $\hun(u)$ on training set $E_0$ for all
$u \in V$ (or setting those values to \shalf{} in case there is no outgoing or incoming edges for
some node), we use the equation \eqref{eq:predictor}\footnote{We reproduce this equation here for
convenience: $\sgn\Big(\big(1-\htr(u)\big) + \big(1-\hun(v)\big) -\tau - \tfrac{1}{2}\Big)$.},
having estimated $\tau$ on $E_0$.

@@

Finally, we can also look in more details at the role of $p$ and $q$ by building a 2D histogram of
the testing edges \euv{} based on their coordinate $(p_u, q_v)$. For instance on
\autoref{fig:tr_synth_mae}, we see that the MAE between $\eta$ and $\wh{\eta}$ is not uniform over
the $p,q$-space. It seems to decrease along the diagonal from $(0,0)$ to $(1,1)$. Note though that
\autoref{fig:tr_synth_sizemae} suggests this is likely due to the higher number of edges in the top
right corner (reflecting the imbalance toward positive edges), which allows better estimation. 
In this case, we cannot use the MCC because outside the diagonal, the Bayes predictor classifies edges
either all positive or all negative, resulting in a division by zero in the definition of MCC. We thus
fall back on accuracy and display in \autoref{fig:tr_synth_acc} the difference between the accuracy
of of \usrule{} and the accuracy of the Bayes predictor. This time, the anti diagonal where
$\frac{p+q}{2}=\lhalf$ seems to play a special role. Indeed, as showed more clearly in
\autoref{fig:tr_synth_cmpacc}, the gap in accuracy between \usrule{} and the Bayes predictor
increases symmetrically as $\frac{p+q}{2}$ moves away from \shalf{}. The gap also tends to disappear
when $\frac{p+q}{2}=\lhalf$, for in that region, both predictors can only rely on random predictions.

While we show those patterns on \wik{} with a 40\% training size, they are consistent across
datasets and training size.

@@

  \begin{subfigure}[b]{0.504\textwidth} 
    \includegraphics[width=\textwidth]{raw/synthgen_wik_rule_40_acc.pdf} 
    \caption{100 time the difference of accuracy between \usrule{} and Bayes.} 
    \label{fig:tr_synth_acc}
  \end{subfigure}~
  \begin{subfigure}[b]{0.504\textwidth} 
    \includegraphics[width=\textwidth]{raw/synthgen_wik_rule_40_cmpacc.pdf} 
    \caption{Difference of accuracy between \usrule{} and Bayes as $\frac{p+q}{2}$ moves away from $\frac{1}{2}$.} 
    \label{fig:tr_synth_cmpacc}
  \end{subfigure}
  \caption{Results on \wik{} with \usrule{} on a training set of size 40\%}
\end{figure*}

@@

  \item The \usrule{} algorithm described in \autoref{ss:passive}.

@@

Our main results are summarized in \autoref{tab:all_mcc}, reporting MCC test set performance after
training on sets of varying size (from 5\% to 25\%, plus 50\% and 90\%). Results have been averaged
over 12 repetitions.
Because scalability is a major concern when training on sizeable datasets, we also give an idea of
relative training times (in milliseconds) by reporting separately in \autoref{tab:all_time} the time it took to train a single run of
each algorithm on a training set of size\footnote{Comparison of training time performances is fair
since all algorithms have been carefully implemented using the same stack of Python libraries, and
run on the same machine (16 Xeon cores and 192Gb Ram).} 15\% of $|E|$, and then predict on the test
set. Some conclusions emerge from those experiments:

@@

  \item \uslpropGsec{} almost always ranks first or at least second in this comparison when MCC is
    considered, at least in the small training set regime (that is when having access to at most
    $25\%$ of the labels, which might be more realistic in many real world situations).
    On top of it, \uslpropGsec{} is fastest among the global methods (one or even two
    orders of magnitude faster), thereby showing the benefit of our approach to \esp{}.

  \item Two methods are competitive with us, \compbayesian{} and \compranknodes{}. \compbayesian{}
    achieves its best results on the larger dataset (\sla{}, \epi{} and \kiw{}) when the training
    set size is large. Indeed, it learns a much more complex model than our methods, with $256$
    parameters\footnote{corresponding to $16\times 16$ feature per edge}, which also requires a lot
    of time to be trained. \compranknodes{} also shines in those large datasets, even at small
    training size. Note however that the difference with \uslpropGsec{} is usually less than one point,
    at the expense of a five times longer learning time.

@@

\input{all_time}

\subsection{Additional experiments}

@@

In the second set of experiments, we study the effect of our hypothesis that the labeled edges of the training set are
chosen \uar{}. In two of our datasets (\wik{} and \epi{}), edges come with the timestamp at which they
were created. A more realistic way of choosing the training set is therefore, for a given training
size $m$, to let $E_0$ be the $m$ oldest edges, and try to predict the remaining, newest ones. This
is a common experimental setting in link prediction, where the goal is to infer which pair of nodes
will be connected in the future~\autocite{linkPredSurvey16}. However, as showed in
\autoref{tab:troll_ts_mcc}, this makes the problem much more challenging, both for our methods and
our competitors. It is quite surprising that even when the training set is as large as 90\% of all
edges, the MCC are so low compared with those reported in \autoref{tab:all_mcc}.

@@

This conclusion is not in contradiction with our previous experiments, since our algorithms were
developed under the assumption that both the revealed labels and the ones to predict were
distributed uniformly within the graph. Recall for instance that \autoref{t:active} requires to have
$\Omega(\log |V|)$ labels outgoing from $u$ and incoming to $v$ to guarantee an accurate prediction
of the sign of \euv{}. Therefore in the next section, we present an online algorithm, whose goal is
to guarantee good performance when the labeling is regular enough, despite the adversarial order in
which the predictions are asked.

@@

### 2017-10-03 troll/related.tex
corrections in troll related and experiments.

\begin{newcontent}
Interest in signed networks can be traced back to the psychological theory of structural
balance~\autocites{Cartwright56}{HeiderBook58} and its weak version~\autocite{davis1967clustering},
that we will describe with more details in \autoref{sub:bias_balance} \vpageref{sub:bias_balance}.
The advent of online signed social networks has enabled a
more thorough and quantitative understanding of that phenomenon. In this section, we provide an
overview of methods tackling the same \esp{} problem as us. Along the way, we give five of them a
name in small capitals, for they are recent and effective. Therefore, we will compare our approaches
with those methods in the experiments of the next section. At the end, we mention some closely
related variants of the original \esp{} problem.

@@

A sophisticated take on degree features is presented by \textcite{Bayesian15}, who note that a node
can belong to one of the 16 types based of whether the number of its positive (respectively
negative) outgoing (respectively incoming) edges is zero or not. The number of unobserved incoming
and outgoing edges of each node $u$ let us define a 16-dimensional vector $V_u$ containing the
probability of transitioning to any other type. Then for each edge \euv, they compute the outer
product of $V_u$ with $V_v$ and also include additional features such as triads count and degree
information before training a Logistic Regression model. We refer to this method as
\emph{\compbayesian{}}.

@@

Other works have also\Todo{maybe that part would be better in the conclusion, especially since side
information would help with the cold start problem showed in the experiments and trust version is
one of the future direction.}
considered versions of the problem where side information related to the network is available to
the learning system. For instance, \autocite{EdgeSignsRating15} uses the product purchased on
\epi{} in conjunction with a neural network, \autocite{TrollDetection15} identifies trolls by
analysing the textual content of their post, and~\autocite{SNTransfer13} uses SVM to perform
transfer learning from one network to another.
\autocite{Tang2013} uses a matrix completion approach, approximating $A$ by $PCP^T$, where the
row $P_u$ of the matrix $P\in\Rbb^{n\times d}$ is the low rank representation of the node $u$ and
$C\in\Rbb^{d\times d}$ is the correlation matrix between these representations. They also assume
they are given a symmetric matrix $Z$ of the homophily coefficients $\zeta_{u,v}$ between $u$ and
$v$ and add the following regularization term to be minimized: $\sum_{u<v} \zeta_{u,v}||P_u -
P_v||_2^2 = Tr(U^T L_Z U)$, where $L_Z$ is the Laplacian of $Z$.
While many of these approaches have interesting
performances, they often require extra information which is not always available (or reliable) and,
in addition, may face severe scaling issues.

@@

While we presented methods that operate in the batch setting, and will present an online algorithm
as well in \autoref{s:algonline}, other works have attacked the \esp{} problem from an active
learning point of view~\autocites{Cesa-Bianchi2012a}{Cesa-Bianchi2012b}.  Recall that in the active
setting, we are given a budget of edge labels to observe, and are free to select them the way we
want within $E(Y)$. Again, the goal is to make as few mistakes as possible when predicting the sign
of the remaining edges. These two methods build a spanning subgraph of $G$ and query all its edge,
but they differ in its the construction. \Textcite{Cesa-Bianchi2012a} partition $G$ into stars and
connect them in a tree, whereas \textcite{Cesa-Bianchi2012b} cover the graph with cycles, each
containing one test edge and being queried for the other, while respecting the user specified query
budget.  This relies on a different bias than our generative model, and this will be the subject of
our next chapter.  One might also consider an online version of the problem where the topology is
not known in advance but discovered as prediction are made~\autocite{Gentile2013}. This naturally
increases and the difficulty of the problem, as reflected by the computational cost of the solution
proposed, which is quadratic in $|V|$ \emph{at each prediction}.
\end{newcontent}

>>>>

Interest in signed networks can be traced back to the psychological theory of structural
balance~\autocites{Cartwright56}{HeiderBook58} and its weak version~\autocite{davis1967clustering},
that we will describe with more details in \autoref{sub:bias_balance} \vpageref{sub:bias_balance}.
The advent of online signed social networks has enabled a
more thorough and quantitative understanding of that phenomenon. In this section, we provide an
overview of methods tackling the same \esp{} problem as us. Along the way, we give five of them a
name in small capitals, for they are recent and effective. Therefore, we will compare our approaches
with those methods in the experiments of the next section. At the end, we mention some closely
related variants of the original \esp{} problem.

@@

A sophisticated take on degree features is presented by \textcite{Bayesian15}, who note that a node
can belong to one of the 16 node-types based of whether the number of its positive (respectively
negative) outgoing (respectively incoming) edges is zero or not. The number of unobserved incoming
and outgoing edges of each node $u$ let us define a 16-dimensional vector $V_u$ containing the
probability of transitioning to any other type once the unobserved signs are revealed. Then each
edge \euv is associated with a feature vector consisting of the outer
product of $V_u$ with $V_v$ and also include additional features such as triads count and degree
information before training a Logistic Regression model. We refer to this method as
\emph{\compbayesian{}}.

@@

Other works have also\done\Todo{maybe that part would be better in the conclusion, especially since side
information would help with the cold start problem showed in the experiments and trust version is
one of the future direction.}
considered versions of the problem where side information related to the network is available to
the learning system. For instance, \autocite{EdgeSignsRating15} uses the product purchased on
\epi{} in conjunction with a neural network, \autocite{TrollDetection15} identifies trolls by
analysing the textual content of their post, and~\autocite{SNTransfer13} uses SVM to perform
transfer learning from one network to another.
\autocite{Tang2013} uses a matrix completion approach, approximating $A$ by $PCP^T$, where the
row $P_u$ of the matrix $P\in\Rbb^{n\times d}$ is the low rank representation of the node $u$ and
$C\in\Rbb^{d\times d}$ is the correlation matrix between these representations. They also assume
they are given a symmetric matrix $Z$ of the homophily coefficients $\zeta_{u,v}$ between $u$ and
$v$ and add the following regularization term to be minimized: $\sum_{u<v} \zeta_{u,v}||P_u -
P_v||_2^2 = Tr(U^T L_Z U)$, where $L_Z$ is the Laplacian of $Z$.
While many of these approaches have interesting
performances, they often require extra information which is not always available (or reliable) and,
in addition, may face severe scaling issues.

@@

While we presented methods that operate in the batch setting, and will present an online algorithm
as well in \autoref{s:algonline}, other works have attacked the \esp{} problem from an active
learning point of view~\autocites{Cesa-Bianchi2012a}{Cesa-Bianchi2012b}.  Recall that in the active
setting, we are given a budget of edge labels to observe, and are free to select them the way we
want within $E(Y)$. Again, the goal is to make as few mistakes as possible when predicting the sign
of the remaining edges. These two methods build a spanning subgraph of $G$ and query all its edge,
but they differ in its the construction. \Textcite{Cesa-Bianchi2012a} partition $G$ into stars and
connect them in a tree, whereas \textcite{Cesa-Bianchi2012b} cover the graph with cycles, each
containing one test edge and being queried for the other, while respecting the user specified query
budget.  This relies on a different bias than our generative model, and this will be the subject of
our next chapter.  One might also consider an online version of the problem where the topology is
not known in advance but discovered as prediction are made~\autocite{Gentile2013}. This naturally
increases and the difficulty of the problem, as reflected by the computational cost of the solution
proposed, which is quadratic in $|V|$ \emph{at each prediction}.
